---
title: 'R code for the manucript "Reliability of citizen scientists’ identification
  features"'
  of camera-trapping images: evaluating misidentification rates and the role environmental
author: "Filipe Rocha"
date: "11/03/2020"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


### Load the Required Packages

```{r}
library(readxl)
library(readr)
library(plyr)
library(Amelia)
library(gdata)
require(multcomp)#compara?oes multiplas
require(lmerTest)#p values em glmm
library(lme4)
library(MASS)
library(car)
library(MuMIn)
library(AICcmodavg)
library(gplots)
library(ggplot2)
library(ggstance)
library(dplyr)
library(stats)

```

### Set the working directory and import the data sets
```{r}

setwd("~/GitHub/CIBIO_Angola/CitizenScience/")
cs_errors <- read.csv("~/GitHub/CIBIO_Angola/CitizenScience/data.csv", sep = ";") %>%
  select(c('error','vegetation','time','species')) %>%
  droplevels(exclude = c("Aardvark","Hare","Jackal","Pangolin","Reptile","Rodent","Vulture","Wild Cat"))

cs_errors$species <- relevel(cs_errors$species,"Elephant")

#data cleaning
sapply(cs_errors, function(x) sum(is.na(x)))

sapply(cs_errors, function(x) length(unique(x)))

missmap(cs_errors, main = "Missing values vs observed")

```


# Modelling the Presence/Absence of Mistakes  

###There are no missing values in the dataset. So I don't need to discard any variable. Although image_id isn't needed for the analisys, so

```{r}

#models

full.model <- glm(error ~ species*time*vegetation, family = binomial(link = 'logit'), data = cs_errors, na.action = "na.fail")
 
model.dredge <- dredge(full.model, beta = FALSE, evaluate = TRUE, rank = "AIC")
model.dredge

model <- glm(error ~ species + time + vegetation + species*time + time*vegetation, family = binomial(link = 'logit'), data = cs_errors, na.action = "na.fail")
(sum.model <- summary(model))


# Analisys of the table of deviance
anova(model, test = "Chisq")   

library(pscl)
pR2(model)  

```

# Plot model estimate
```{r}
beta <- sum.model$coefficients
CI <- confint(model)
L <- CI[,1]
U <- CI[,2]
names <- c("(Intercept)", "Baboon", "Bird (other)", "Buffalo", "Bushbuck", "Bushpig", "Civet", "Duiker", "Genet", "Hippopotamus",
           "Honey Badger", "Impala", "Kudu", "Mongoose", "Nyala", "Oribi", "Porcupine", "Reedbuck", "Sable Antelope", "Samango Monkey",
           "Serval", "Vervet Monkey", "Warthog", "Waterbuck", "night", "Limestone Gorge", "Miombo Woodland", 
           "Mixed Savanna and Woodland", "Baboon*night", "Bird (other)*night", "Buffalo*night", "Bushbuck*night", "Bushpig*night",
           "Civet*night", "Duiker*night", "Genet*night", "Hippopotamus*night", "Honey Badger*night", "Impala*night", "Kudu*night",
           "Mongoose*night", "Nyala*night", "Oribi*night", "Porcupine*night", "Reedbuck*night", "Sable Antelope*night","Serval*night",
           "Warthog*night", "Waterbuck*night", "night*Limestone Gorge", "night*Miombo Woodland", "night*Mixed Savanna and Woodland")

L <- L[-c(47, 49)]
U <- U[-c(47, 49)]

beta.ci <- data.frame(names,beta,L,U)
beta.ci <- beta.ci[-c(1),]

beta.ci <- subset(beta.ci, Pr...z..< 0.05)
beta <- beta.ci$Estimate
L <- beta.ci$L
U <- beta.ci$U

plotCI(beta, ui=U, li=L)


```






# Number of Specimens Misidentified
```{r}

cs_nerrors<- read.csv("~/GitHub/CIBIO_Angola/CitizenScience/data.csv", header = T, sep = ";") %>%
  select(c(5,8,9,12:16)) %>%
  droplevels(exclude = c("Aardvark","Hare","Jackal","Pangolin","Reptile","Rodent","Vulture","Wild Cat")) %>%
  rename(log.sum_n_error = log.sum_error..1) %>%
  mutate(vegetation = as.factor(vegetation),
         time = as.factor(time),
         species = as.factor(species),
         greg = as.factor(greg),
         log.sum_n_error = as.double(log.sum_n_error))

cs_nerrors$species <- relevel(cs_nerrors$species,"Bushbuck")
cs_nerrors$greg <- relevel(cs_nerrors$greg,"0")

#Data cleaning
sapply(cs_nerrors, function(x) sum(is.na(x)))
cs_nerrors <- na.omit(cs_nerrors)
sapply(cs_nerrors, function(x) sum(is.na(x)))

sapply(cs_nerrors, function(x) length(unique(x)))

missmap(cs_nerrors, main = "Missing values vs observed")

```



### models
```{r}


full.glm <- glm(log.sum_n_error ~ species*time*vegetation*greg, family = gaussian, data = cs_nerrors, na.action = "na.fail")
summary(full.glm)
# out <- capture.output(summary(full.glm))
# cat("full.glm", out, file="summary_full.glm.txt", sep="\n", append=FALSE)


full.glm.dredge <- dredge(full.glm, beta = FALSE, evaluate = TRUE, rank = "AIC")
full.glm.dredge

write.table(x = full.glm.dredge, file = "full.glm.dredge.txt", sep = "\t")

(avg.glm <- model.avg(full.glm.dredge, beta=TRUE, rank=AIC, fit = TRUE)) ##faz o "modelo medio" usado valores de delta iguais ou inf
(sum.avg.glm <- summary(avg.glm))

(CI.glm <- confint(avg.glm))


```


###  Plot Estimates
```{r}

library(tidyr)

 beta.glm <- read.csv("beta_GLM.csv", sep = ";") %>%
   mutate(Estimate = as.numeric(Estimate),
          CI2.5 = as.numeric(CI2.5),
          CI97.5 = as.numeric(CI97.5))

ls.str(beta.glm)


(k <- ggplot(beta.glm, aes(x = factor(Interact, levels = unique(Interact) ), y = Estimate, abline(y = 1.0), group = Model, color = Model)) +  
     geom_hline(yintercept = 0) +
     geom_pointrange(aes(ymin = Estimate-CI2.5, ymax = Estimate+CI97.5), position = position_dodge(width = 0.6), na.rm=TRUE) +
     theme_classic(base_size = 11, base_family = "") +
     ylab(expression(beta.glm)) + theme(legend.position = "bottom") + 
     theme(axis.text.x = element_text(angle = 90, vjust=0.5)))

```


### Tornado Chart: Number of errors (False Positives & Negatives)
```{r}

library(plotrix)

tornado <- read.csv("tornado.csv", sep = ";")

falneg <- tornado$perc.fn
fal_pos <- tornado$perc.fp
splabels <- tornado$Species
fncol <- gray.colors(start = 0.6, end = 0.6, n = 1)
fpcol <- gray.colors(start = 0.3, end = 0.3, n = 1)
par(mar=pyramid.plot(falneg,fal_pos,labels=splabels,
                     top.labels = c("False Negatives", "Species", "False Positives"), xlim = c(1.0, 1.0), 
                     laxlab = seq(from = 0, to = 1.0, by = 0.25),
                     raxlab=seq(from = 0, to = 1.0, by = 0.25),  
                     lxcol=fncol,rxcol=fpcol,
                     gap=0.2, show.values=FALSE))

```


### ChordDiagram: Plot the proportion of records wrongly assigned to a species, and the proporsion of the most frequent confusions
```{r}

herbivores <- read.table("herbivores.txt", skip=2, stringsAsFactors=FALSE)

#select and rename labels
library("dplyr")
data <- herbivores %>% select(1:3) %>% 
  rename(order = V1, 
         rgb = V2, 
         species = V3) %>% 
  mutate(species = gsub("_", " ", species))

#flow matrix
m <- as.matrix(herbivores[,-(1:3)]/1e06)
dimnames(m) <- list(orig = data$species, dest = data$species)
#drop small flows
m[m<=quantile(m,0.65)]<-0

#sort species and create colours
library("tidyr")
data <- data %>% 
  arrange(order) %>% 
  separate(rgb, c("r","g","b")) %>% 
  mutate(col = rgb(r, g, b, max=255), 
         max = rowSums(m)+colSums(m))

#plot using chordDiagram
library("circlize")
circos.clear()
par(mar = rep(0, 4), cex=0.9)
circos.par(start.degree = 90, gap.degree = 4)
chordDiagram(x = m, directional = 1, direction.type = c("arrows"), order = data$species, 
             grid.col = data$col, annotationTrack = "grid", 
             transparency = 0.25,  annotationTrackHeight = c(0.1, 0.1),
             diffHeight  = -0.05)


#add in labels and axis
circos.trackPlotRegion(track.index = 1, panel.fun = function(x, y) {
  xlim = get.cell.meta.data("xlim")
  sector.index = get.cell.meta.data("sector.index")
  circos.text(mean(xlim), 2.5, sector.index, facing = "bending")
  circos.axis("top", major.at = seq(0, max(xlim)), minor.ticks=1, labels.away.percentage = 0.2, labels.niceFacing = FALSE )
}, bg.border = NA)
circos.clear()


```


# Code for the estimation of the required number of records to accuratly know the crowdsourcing misidentification rate

### Organize the data for the analysis.
```{r}

allspecies <- cs_errors <- read.csv("~/GitHub/CIBIO_Angola/CitizenScience/data.csv", sep = ";") %>%
  select(c("image_id", "error", "false_pos", "false_neg",  "species")) %>%
  mutate(n = (1:length(image_id))) %>%
  filter(species != "Aardvark",          
         species != "Eland",
         species != "Ground Hornbill",
         species != "Pangolin",
         species != "Hare",
         species != "Hartebeest",
         species != "Jackal",
         species != "Lion",
         species != "Raptor (other)",
         species != "Reptile",
         species != "Rodent",
         species != "Vulture",
         species != "Wild Cat",
         species != "Wildebeest",
         species != "Zebra")

max <- length(allspecies$n)
rmax <- 1000 
i <- as.integer()
j <- as.integer()
n2 <- seq.int(nrow(allspecies))
int <- data.frame(n2)
erros <- matrix(NA,nrow(allspecies),ncol = rmax)
perc <- matrix(NA,nrow(allspecies),ncol = rmax)
final <- data.frame(n2)
names <- vector(,length = rmax)
regression <- data.frame(matrix(vector(), 0, 11,
                                dimnames=list(c(), c("Rand", "Threshold_Y=0", "perc_Y=0", "B_Y=0", "A_Y=0","p_Y=0", 
                                                     "Threshold_nonsig", "perc_nonsig", "B_nonsig", "A_nonsig","p_nonsig"))),
                         stringsAsFactors = F)
output <- data.frame(matrix(vector(), 0, 5,
                             dimnames=list(c(), c("n", "B", "A", "AbsoluteA","p"))),
                      stringsAsFactors = F)

sps <- as.character(unique(allspecies$species))

for (i in 1:length(sps)+1) {
  ifelse(i <= length(sps), species <- subset(allspecies, allspecies$species == sps[i]), species <- allspecies) 


  ## Randomize the data frame
  
  for(x in 1:rmax){
    
    print(x)
    
    set.seed(x)
    rand <- sample(nrow(species))  #creates a random sequence of numbers from 1 to nrow
    
    species <- species[rand, ]      # order the data frame according to random order defined in rand
    
    
    #Create new colums for error and percents
    erros[1,x] <- species$false_pos[1]+species$false_neg[1]
    
    for(i in 2:max){
      j <- i - 1
      erros[i,x] <- erros[j,x] + species$false_pos[i] + species$false_neg[i]
      #    print(c(x, i, errors[i,x]))
    }
    int <- cbind(int,erros[,x])
    int <- transform(int, perc = erros[,x] / n2)
    perc[,x] <- int$perc
    final <- cbind(final,perc[,x])
  }
  
  
  ## Linear Regression

  for(x in 501:600){
    for(i in 1:(max-1)){
      final1 <- subset(final, n2>= i)
      
      bbmodeln <- lm(final1[ ,(x+1)]~n2, data = final1)
      coeff <- coefficients((bbmodeln))
      
      output[i,1] <- i
      output[i,2] <- coeff[1]
      output[i,3] <- coeff[2]
      output[i,4] <- abs(coeff[2])
      output[i,5] <- summary(bbmodeln)$coefficients[2,4]
      
    }
    print(x)
    
    

    ## Non significant

    
    test <- 0
    
    if (nrow(output)>0){
      for(y in 1:(length(output$n)-1)){
        if(output[y,5] > 0.05){
          test <- test + 1
        } else{
          test <- test + 0
        } 
      }
    }
    
    if(test == 0){
      regression[x,1] <- x
      regression[x,2] <- NA
      regression[x,3] <- NA
      regression[x,4] <- NA 
      regression[x,5] <- NA
      regression[x,6] <- NA
      regression[x,7] <- NA
      regression[x,8] <- NA 
      regression[x,9] <- NA
      regression[x,10] <- NA
      regression[x,11] <- NA
    }
    else{
      
      
      sig <- subset(output, p > 0.05)
      sig <- sig[order(sig$n),]
      sig_t <- sig[1,1]
      
      sig_data <- subset(final, n2 >= sig_t)
      
      model.p <- lm(sig_data[,x+1]~n2, data = sig_data)
      
      

      ### Crossing Y = 0

      
      sort_output <- output[order(output$AbsoluteA),]
      cross_t <- sort_output[1,1]
      
      cross0_data <- subset(final, n2 >= cross_t)
      
      model.0 <- lm(cross0_data[,x+1]~n2, data = cross0_data)
      
      
      ### Paste results in the table
      
      regression[x,1] <- x
      regression[x,2] <- cross_t
      regression[x,3] <- cross_t*100/max
      regression[x,4] <- summary(model.0)$coefficients[2,4]
      regression[x,5] <- summary(model.0)$coefficients[1,1]
      regression[x,6] <- summary(model.0)$coefficients[2,1]
      regression[x,7] <- sig_t
      regression[x,8] <- sig_t*100/max
      regression[x,9] <- summary(model.p)$coefficients[2,4]
      regression[x,10] <- summary(model.p)$coefficients[1,1]
      regression[x,11] <- summary(model.p)$coefficients[2,1]
    }
  
    regression1 <- regression
      
  }
  

  ## Results Analysis
  
  regression <- regression[complete.cases(regression), ]
  
 
  ### Histogram 

  
  hist_breaks <- seq(1,length(species$false_neg),1)
  
  hist(regression$Threshold_Y.0, breaks = hist_breaks)
  hist(regression$Threshold_nonsig, breaks = hist_breaks)
  
  mean(regression$Threshold_Y.0)
  
  
  bs <- function(formula, data, b) {
    d <- data[b,] # allows boot to select sample 
    fit <- lm(Threshold_nonsig~Rand, data=d)
    return(coef(fit))
  }
  
  
  ## Bootstrapping

  ### Y = 0
  
  library(boot)
  
  bs <- function(formula, data, b) {
    d <- data[b,] # allows boot to select sample 
    fit <- lm(Threshold_Y.0~Rand, data=d)
    return(coef(fit))
  }
  
  #### bootstrapping with 1000 replications 
  results <- boot(data=regression, statistic=bs, 
                  R=1000, formula=Threshold_Y.0~Rand)
  
  #### view results
  results
  plot(results, index=1) # intercept 
  plot(results, index=2) # slope 
  
  
  #### get 95% confidence intervals 
  boot.ci(results, type="bca", index=1) # intercept 
  boot.ci(results, type="bca", index=2) # slope
  

  ### p > 0.05

  
  
  bs <- function(formula, data, b) {
    d <- data[b,] # allows boot to select sample 
    fit <- lm(Threshold_nonsig~Rand, data=d)
    return(coef(fit))
  }
  
  #### bootstrapping with 1000 replications 
  results <- boot(data=regression, statistic=bs, 
                  R=1000, formula=Threshold_nonsig~Rand)
  
  #### view results
  results
  plot(results, index=1) # intercept 
  plot(results, index=2) # slope 
  
  
  #### get 95% confidence intervals 
  boot.ci(results, type="bca", index=1) # intercept 
  boot.ci(results, type="bca", index=2) # slope

}

```


```{r}

ccr <- read.csv("CumulativeCurveResults.csv", sep = ";", header = TRUE) %>%
  filter(Species != "Hartebeest",
         Species != "Wildebeest", 
         Species != "All species") %>%
  mutate(log.n = log(n),
         Species = recode(Species, 'Bird' = 'Bird (other)'))



# GLM 

m1 <- glm(r_T~n, data = ccr, family = gaussian(link = 'identity'))
m2 <- glm(r_T~log(n), data = ccr,family = gaussian(link = 'identity'))

require(MuMIn)
model.sel(m1,m2)
summary(m2)

#prediction
x <- seq(0,max(ccr$n),50) 
x <- data.frame(x)
colnames(x) <- "n"

yf <- data.frame(predict(m2, type="link", newdata=x, re.form=NA, se.fit=T))
colnames(yf) <-c("fit","se")
yf <- yf[-1,]
yf$fit<- yf$fit
yf$inf <- yf$fit-1.96*yf$se
yf$sup <- yf$fit+1.96*yf$se
yf <- cbind(x[-1,],yf)
colnames(yf)[1] <- 'n'


require(ggplot2)
(ccr_plot <- ggplot(data = ccr, aes(x = n, y = r_T, ymax = 0.47)) + geom_point() +
  xlab("Sample Size") + ylab("Relative Threshold")+
  theme_classic())

(ccr.glm_plot <- ggplot(data = ccr, aes(x = n, y = r_T)) + geom_point() + 
  geom_ribbon(data=yf, aes(x= n, y = fit, ymin = inf, ymax = sup)) + 
  geom_line(data = yf, aes(x = n, y = fit)) +
  xlab("Sample Size") + ylab("Relative Threshold")+
  theme_classic())

```

## Test of the extrapolation of a subset created with the threshold
``` {r}

data <- read.csv("~/GitHub/CIBIO_Angola/CitizenScience/data2.csv", sep = ";")

# Clean the data and remove the species that do not enter in the analysis
data.correct <- data %>%
  group_by(species_conf) %>%
  summarize(n = n()) %>% # Count the number of pictures for each species, identified by the expert
  filter(species_conf != "Aardvark",          
         species_conf != "Eland",
         species_conf != "Ground Hornbill",
         species_conf != "Pangolin",
         species_conf != "Hare",
         species_conf != "Hartebeest",
         species_conf != "Jackal",
         species_conf != "Lion",
         species_conf != "Raptor (other)",
         species_conf != "Reptile",
         species_conf != "Rodent",
         species_conf != "Vulture",
         species_conf != "Wild Cat",
         species_conf != "Wildebeest",
         species_conf != "Zebra",
         species_conf != "Vervet Monkey",
         species_conf != "Hippopotamus",
         species_conf != "Honey Badger",
         species_conf != "Human",
         species_conf != "Serval",
         species_conf != "Porcupine")

data.correct.all <- data.correct

## we use a for loop to estimate the number for each species at the time
for (i in 1:(nrow(ccr)-1)) { 
  species <- data %>%
  filter(species_CS == as.character(ccr$Species[i])) %>% # Filter the data to use the target species
  sample_frac(size = ccr$r_T[i]) %>% # Sample the species records, according to the threshold calculated above
  group_by(species_conf)%>%
  summarize(n = n()) %>% 
  mutate(prop = n/sum(n), # calculate the proportion of the pictures, iniatially assigned as the target species, distributed according the correct species
         n_correct = as.integer(prop*ccr$n[i])) %>% # estimation of the number of pictures for each species
  select(c(1,4))
  
  data.correct <- data.correct %>%
    left_join(species, by = "species_conf") # paste the results in a data frame with all the results
   colnames(data.correct)[(i+2)] <- paste(ccr$Species[i])
  
}
  
data.correct <- data.correct %>%
  replace(is.na(.), 0) %>%
  mutate(n.correct = rowSums(.[3:21])) # sum up of the number of pictures estimated for each species

## Estimation using the threshold for all species 
for (i in 1:(nrow(ccr)-1)) {
  species <- data %>%
  sample_frac(size = ccr$r_T[20]) %>% # Sample the species records, according to the threshold calculated for all species together
  group_by(species_conf)%>%
  summarize(n = n()) %>%
  mutate(prop = n/sum(n),  # calculate the proportion of the pictures, iniatially assigned as the target species, distributed according the correct species
         n_correct = as.integer(prop*ccr$n[i])) %>% # estimation of the number of pictures for each species
  select(c(1,4))
  
  data.correct.all <- data.correct.all %>%
    left_join(species, by = "species_conf")  # paste the results in a data frame with all the results
   colnames(data.correct.all)[(i+2)] <- paste(ccr$Species[i])
  
}
  
data.correct.all <- data.correct.all %>%
  replace(is.na(.), 0) %>%
  mutate(n.correct.all = rowSums(.[3:21]))  # sum up of the number of pictures estimated for each species

(final.correct <- data.correct %>%
  left_join(data.correct.all, by = "species_conf") %>% # join both tables with results
  select(c('species_conf','n.x','n.correct', 'n.correct.all')) %>% # select the columns with the sum
  rename('n' = 'n.x') %>%
  mutate(n.diff = abs(n.correct - n),
         n.diff.all = abs(n.correct.all - n)) # Calculate the difference between the real and the estimated number of pictures, using the threshold for each species isolated and for all the species together
)

write.csv(final.correct, file = "data.correction.csv")




```


